For Finitary Induction-Induction, \\ Induction is Enough

### 2020-05-03 paper/paper.tex
finish

We describe related work in \Cref{sec:related}, and explain our
notation and Agda formalisaton in \Cref{sec:notation}. Then the
following three sections describe our three contributions:
\begin{itemize}
\item \Cref{sec:theory_of_signatures}. We define what it means for a
  model of extensional type theory (ETT, \Cref{def:ett}) to support
  all inductive-inductive types (IITs): \Cref{def:hasiits}. The novel
  contribution here is a (predicative) Church encoding of signatures
  following \cite{DBLP:conf/lics/AwodeyFS18}.
\item \Cref{sec:constructingiits}. We show that if a model of ETT
  supports the theory of IIT signatures
  (\Cref{def:theoryofsignatures}), then it supports IITs in the above
  sense: \Cref{thm:ToSToIITs}. This is an adaptation of a proof in
  \cite{Kaposi:2019:CQI:3302515.3290315}.
\item \Cref{sec:ambroise}. Our main contribution is showing that if a
  model of ETT supports indexed W-types, then it supports the theory
  of IIT signatures (\Cref{thm:WToToS}), and hence, all IITs (\Cref{thm:WToIITs}).
\end{itemize}
We conclude in \Cref{sec:conclusions}.

@@

TODO

The current work only concerns finitary IITs. An extension would be to also
allow infinitely branching inductive types such as W-types. This would first
require giving semantics for infinitary IITs (to our knowledge there is no
previously published semantics that we can borrow), and also giving a term model
construction analogously to finitary QIITs. These steps seem feasible. However,
it seems to be more difficult to construct the syntax of infinitary IIT
signatures without using quotients. The reason is that such syntax would not be
strictly restricted to neutral terms: we would need $\lambda$-abstraction and
$\beta\eta$-rules for infinitary $\Pi$ types, in order to allow a term model
construction for infinitary IITs. A definition of normal preterms and typing
judgments on them may still be possible, but it appears to be much more
complicated than before (the current authors have attempted this without
conclusive success).

>>>>

We describe related work in \Cref{sec:related}, and explain our
notation and Agda formalisaton in \Cref{sec:notation}. Then the
following three sections describe our three contributions:
\begin{itemize}
\item \Cref{sec:theory_of_signatures}. We define what it means for a
  model of extensional type theory (ETT, \Cref{def:ett}) to support
  all inductive-inductive types (IITs): \Cref{def:hasiits}. The novel
  contribution here is a (predicative) Church encoding of signatures
  following \cite{DBLP:conf/lics/AwodeyFS18}.
\item \Cref{sec:constructingiits}. We show that if a model of ETT
  supports the theory of IIT signatures
  (\Cref{def:theoryofsignatures}), then it supports IITs in the above
  sense: \Cref{thm:ToSToIITs}. This is an adaptation of a proof in
  \cite{Kaposi:2019:CQI:3302515.3290315}.
\item \Cref{sec:ambroise}. Our main contribution is showing that if a
  model of ETT supports indexed W-types, then it supports the theory
  of IIT signatures (\Cref{thm:WToToS}), and hence, all IITs (\Cref{thm:WToIITs}).
\end{itemize}
We list further work in \Cref{sec:conclusions}.

@@

The current work only concerns finitary IITs. An extension would be to
also allow infinitely branching inductive types such as W-types. This
would first require giving semantics for infinitary IITs and adapting
the term model construction. These would be straightforward following
\cite{large_inf_qiit}. However, it seems to be more difficult to
construct the syntax of infinitary IIT signatures without using
quotients. The reason is that such syntax would not be strictly
restricted to neutral terms: the term model construction for
infinitary IITs requires $\lambda$-abstraction and $\beta\eta$-rules
for infinitary $\Pi$ types. A definition of normal preterms and typing
judgments on them may still be possible, but it appears to be much
more complicated than before (the current authors have attempted this
without conclusive success).
